---
title: "Class11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database for biomolecular structure data

> Q1: Determine the percentage of PDB structures solved by X-Ray and Electron Microscopy. 

> Determine what proportion of structures are protein?

Download csv file from PDB website
```{r}
# Read CSV
data <- read.csv("Data Export Summary.csv", row.names = 1)
data
```

Total number of entries
```{r}
sum(data$Total)
```

Proportion of entries from each method 
```{r}
round ((data$Total/sum(data$Total)) * 100, 2)
```

Proportion that are protein
```{r}
round (sum(data$Proteins)/sum(data$Total), 2)
```

## HIV-Pr structure analysis 

Here we will read the 1HSG PDB structure and select the protein component and write out a new **protein-only** PDB format file. We then do the same for the ligand (i.e. known drug molecule) creating a **ligand-only** PDB file. 
```{r}
library(bio3d)
pdb <- read.pdb("1hsg.pdb")
pdb
```

```{r}
# read.pdb()
# atom.select()
# write.pdb()
# trim.pdb()
```

```{r}
ligand <- atom.select(pdb, "protein", value=TRUE)
write.pdb(ligand, file="1hsg_ligand.pdb")
```

