---
title: "Class09"
author: "Cassandra Wang"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory data analysis
```{r}
# Save your input data file to a new 'data' directory
fna.data <- "WisconsinCancer.csv"
```

```{r}
# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data)
```

```{r}
head(wisc.df)
```

We have `r nrow(wisc.df)` samples in this dataset
```{r}
nrow(wisc.df)
```

How many benign and malignant samples do we have in this dataset?
```{r}
table(wisc.df$diagnosis)
```

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

head(wisc.data)
```

Store the diagnosis for reference in the future as a separate vector
```{r}
diagnosis <- wisc.df$diagnosis
```

Q1: How many observations are in this dataset?Number of patient smamples and observations for each
```{r}
nrow(wisc.data)
```

Q2: How many of the observations have a malignant diagnosis?
```{r}
table(wisc.df$diagnosis)
```

Q3: How many variables/features in the data are suffixed with "_mean"?
```{r}
grep("_mean", colnames(wisc.df), value=TRUE)
```

I can use `length()` to counth how many matches we have
```{r}
length(grep("_mean", colnames(wisc.df)))
```

## Performing Principal Component Analysis
```{r}
# Check column means 
colMeans(wisc.data)
```

```{r}
# Check standard deviations
round(apply(wisc.data,2,sd), 3)
```

These values look very different so I will use `scale=TRUE` when I run PCA

```{r}
# Perform PCA on wisc.data 
wisc.pr <- prcomp(wisc.data, scale=TRUE)

# Look at summary of results
summary(wisc.pr)
```

```{r}
# To get summary results for PC1
x <- summary(wisc.pr)
x$importance[,"PC1"]
```

```{r}
biplot(wisc.pr)
```

```{r}
# Scatter plot observations by components 1 and 2
plot(wisc.pr$x[,1], wisc.pr$x[,2] , col = diagnosis, 
     xlab = "PC1", ylab = "PC2")
```

```{r}
# Scatter plot observations by components 1 and 3
plot(wisc.pr$x[,1], wisc.pr$x[,3] , col = diagnosis, 
     xlab = "PC1", ylab = "PC2")
```

```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
# Variance explained by each principal component: pve
pve <- sum(pr.var)/pr.var

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```


## Hierarchical Clustering
```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.hclust <- hclust(data.dist)
```

```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
```

```{r}
table(wisc.hclust.clusters, diagnosis)
```

## Clustering on PCA results

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model
```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method="ward.D2")

plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

Cluster 1 appears to be mostly malignant, and Cluster 2 appears to be mostly benign. 

```{r}
plot(wisc.pr$x[,1:2], col=grps+1)
```

Turn our groups into a factor and reorder the levels so cluster 2 (more benign) comes first
```{r}
g <- as.factor(grps)
levels(g)

g <- relevel(g,2)
levels(g) 
```

```{r}
# Plot using our re-ordered factor 
plot(wisc.pr$x[,1:2], col=g)
```

Cut the hierarchical clustering model into 2 clusters
```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
```

```{r}
table(wisc.pr.hclust.clusters, diagnosis)
```

## Prediction

Take our PCA model from before and new cancer cell data and project that data onto our PCA space
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

Black cluster is Cluster 1 (benign), red is Cluster 2 (malignant).
We should prioritize Patient 2 for follow-up. 

