---
title: "Class08"
author: "Cassandra Wang"
date: "10/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering
```{r}
# Generate some sample data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

Use the kmeans() function setting k to 2 and nstart=20
```{r}
k <- kmeans(x, centers=2, nstart=20)
```

```{r}
k
```

For cluster size:
```{r}
k$size 
```

For cluster membership/assignment:
```{r}
k$cluster
```

For cluster center:
```{r}
k$centers
```

Plot x colored by the kmeans cluster assignment and add cluster centers as blue points
```{r}
plot(x, col=k$cluster)
points(k$centers, col="blue", pch=15)
```

## Hierarchical clustering
The `hclust()` function requires a distance matrix as an input. You can get this from the `dist()` function. 

```{r}
# First we need to calculate point (dis)similarity as the Euclidean distance between observations
dist_matrix <- dist(x)

# The hclust() function returns a hierarchical clustering model
hc <- hclust(d = dist_matrix)

# The print method is not so useful here
hc
```

```{r}
# Draw a dendrogram
plot(hc)
abline(h=6, col="red")
grps <- cutree(hc, h=6)
```

```{r}
cutree(hc, k=4)
```

```{r}
# Step 1. Generate some example data for clustering
x2 <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x2) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x2)

# Step 3. Generate colors for known clusters (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x2, col=col)
```

```{r}
# clustering 
hc2 <- hclust(dist(x2))

# draw tree
plot(hc2)
abline(h=2, col="red")

# cut the tree into clusters/groups 
grps2 <- cutree(hc2, k=3)
grps2
```

Plot the data colored by their hclust result with k=3. 
```{r}
plot(x2, col=grps2)
```

How many points in each cluster?
```{r}
table(grps2)
```

Cross-tabulate (i.e. compare our clustering result with the known answer)
```{r}
table(grps2, col)
```

## Principal Component Analysis (PCA)

Read some example gene expression data: 
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)

head(mydata)
```

How many genes are is this dataset?
```{r}
dim(mydata)
nrow(mydata)
```

Let's do pca with the pr(comp) function.
```{r}
pca <- prcomp(t(mydata), scale=TRUE)

## See what is returned by the prcomp() function
attributes(pca)

# $names
#[1] "sdev" "rotation" "center" "scale" "x"
#
# $class
#[1] "prcomp"

## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2])

## Variance captured per PC
pca.var <- pca$sdev^2 

## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per

barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")

## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```

## PCA of UK Food Data
```{r}
x <- read.csv("UK_foods.csv")
dim(x)
```

Preview first and last 6 rows of the dataset. 
```{r}
head(x)
tail(x)
```

```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

```{r}
dim(x)
```

```{r}
x <- read.csv("UK_foods.csv", row.names=1)
head(x)
```

Generating barplots and pairwise plots
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

# Use the prcomp() PCA function 
```{r}
pca <- prcomp( t(x) )
summary(pca)
```

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

```{r}
plot(pca$x[,1], pca$x[,2])
```

