---
title: "Class07"
author: "Cassandra Wang"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R functions revisited

Source my functions from last class:

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```

```{r}
rescale( c(1, 10, 5, NA, 6))
```

```{r}
rescale2( c(1:10, "name"))
```

## A new function called both_na()
Write a function to find where there are NA elements in two input vectors:

First, make some simple input where I know the answer. 
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

Look online to found the **is.na()** function.
```{r}
is.na(x)
is.na(y)
```

The **which()** function tells me where the TRUE values are.
```{r}
which( is.na(x))
```

The AND function requires two input TRUE to give a TRUE output.
```{r}
is.na(x) & is.na(y)
```

Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have. This is my working snippet of code. 
```{r}
sum( is.na(x) & is.na(y))
```

Now turn it into a function.
```{r}
both_na <- function(x,y) {
  sum( is.na(x) & is.na(y))
}
```

Testing my function.
```{r}
both_na(x,y)
```

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x, y2)
```

```{r}
x2 <- c(NA,NA)
both_na(x2, y2)
```

Demonstrate re-cycling of input. 
```{r}
x <- c(1, NA, NA)
y3 <- c( 1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

```{r}
length(x)
length(y3)
```

Add a check for when inputs x and y are not the same length. 
```{r}
both_na2 <- function(x,y) {
  if(length(x) != length(y)) {
    stop("Inputs x and y should be the same length!")
  }
  sum( is.na(x) & is.na(y))
}
```

```{r}
both_na2(x,y3)
```

## Writing a new function grade():

Working code for our function
```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
mean(student1[-which.min(student1)])
mean(student2[-which.min(student2)], na.rm=TRUE)
```

Turn our working code into a first function
```{r}
grade <- function(x) {
  if( any(is.na(x)) ){
    warning("Student is missing a homework")
  }
  mean( x[-which.min(x)], na.rm=TRUE)
}
```

For students that have multiple NA grades 
```{r}
grade2 <- function(x) {
  if( any(is.na(x)) ){
    warning("Student is missing a homework")
  }
  sum( x[-which.min(x)], na.rm=TRUE)/length(x)-1
}
```

Testing our function
```{r}
grade(student1)
grade(student2)

grade2(student2)
```

```{r}
url <- ("https://tinyurl.com/gradeinput")
hw <- read.csv(url, row.names=1)
```

```{r}
apply(hw, 1, grade)
```

